---
title: "conditionalsRSA-rwebppl"
author: "Britta Grusdt"
date: "2/3/2020"
output: html_document
---

```{r setup, include=FALSE}
library(forcats)
library(ggplot2)
library(ggthemes)
library(here)
library(rwebppl)
library(tidyverse)
source(here("utils.R"))
source(here("experiments", "blocksworld-main", "analysis", "model-utils.R"))
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
TARGET_DIR <- here("experiments", "blocksworld-main", "results", "model")
dir.create(TARGET_DIR, recursive = TRUE, showWarnings = FALSE)
```

# Behavioral data from experiments

```{r eval=FALSE, message=FALSE}
fn <- here("experiments", "blocksworld-prior", "results", "data-processed",
           "tables_experimental.csv")
dat.experiment1 <- read_csv(fn)

fn <- here("experiments", "blocksworld-main", "results", "data-processed", 
           "data_experimental_means.csv")
dat.experiment2.means <- read_csv(fn) %>%
  mutate(utterance=factor(utterance)) %>%
  add_column(produced_by = "human")


# dat.experiment2.normalized <- dat.experiment2.means %>%
#   group_by(stimulus_id) %>%
#   mutate(mean=mean/sum(mean))
# stopifnot(dat.experiment2.normalized %>% filter(is.infinite(mean)) %>% nrow()==0);
```

# setup to run model

```{r}
dat.to_wppl <- list()
tables_to_wppl <- prepare_tables(dat.experiment1)

dat.to_wppl$stimulus_ids = tables_to_wppl %>% pull(stimulus_id) %>% unique()
dat.to_wppl$tables <- tables_to_wppl

# check for utterances with highest theta
# (are there any that are not covered by states?)
dat.utts <- pids_to_cover_utts(dat.experiment1, 0.9)
stopifnot(dat.utts$utts_not_covered %>% length == 0);
dat.to_wppl$utterances <- dat.utts$utts_covered
# dat.to_wppl$par_dirichlet = read_csv("results-dirichlet.csv") %>%
#                           add_column(alpha=list(alpha_1, alpha_2, alpha_3, alpha_4))
#                       nest("alpha_1", "alpha_2", "alpha_3", "alpha_4")  %>%
#                       select(stimulus_id, alpha) %>%
#                       pivot_wider(names_from = stimulus_id, values_from = alpha) %>% unnest()

```

# RSA model
Mean squared Errors over grid

```{r, echo=FALSE, message=FALSE, warning=FALSE}
alphas <- c(7)
thetas <- c(0, 0.1, 0.7, 0.8)
cost_conditional <- c(0.25, 0.4)
cost_other <- c(5, 8, 10)
sweep <- expand.grid(alpha=alphas, theta=thetas, cost_if=cost_conditional,
                     cost_other=cost_other) %>% transpose() 

now <- str_replace_all(Sys.time(), "[\\s+:-]", "_")
target_dir <- paste(TARGET_DIR, "mse-grid", sep=.Platform$file.sep)
dir.create(target_dir, recursive = TRUE, showWarnings = FALSE)

for(idx_run in seq(1, length(sweep), by = 1)) {
  if(idx_run %% 5 == 0 ){
    print(paste('starting run', idx_run))
  }
  dat.to_wppl$alpha <- sweep[[idx_run]]$alpha
  dat.to_wppl$theta <- sweep[[idx_run]]$theta
  dat.to_wppl$cost_conditional <- sweep[[idx_run]]$cost_if
  dat.to_wppl$cost_and <- sweep[[idx_run]]$cost_other
  dat.to_wppl$cost_likely <- sweep[[idx_run]]$cost_other
  dat.to_wppl$cost_neg <- sweep[[idx_run]]$cost_other
  
  model_predictions <- run_model(dat.to_wppl)
  model <- model_predictions$avg
  # use UNnormalized participant ratings for MSE!
  mse <- mse_model_human(dat.experiment2.means, model)
  
  save_predictions(model %>% add_column(run=idx_run),
                   target_dir, "results_model", append_to_file=TRUE)
  save_params(dat.to_wppl, target_dir, idx_run, mse, append_to_file=TRUE)
}
```

## Notes after several runs

- The higher the cost for other utterances than conditional, the better; it
shouldn't be 0. Around 0.25 seems good

- theta around 0.6-0.7 seems to be good, but still good results for very low
theta of 0.25 and even for theta = 0

- best result with theta=0.6 with cost_conditional=0.1

- alpha doesn't seem to have a large influence on results


# Look at results of parameter grid

```{r, message=FALSE}
params <- read_csv(paste(target_dir,"parameters.csv",sep=.Platform$file.sep))%>%
  arrange(MSE) %>% select(-run) %>%
  mutate(alpha=factor(alpha), theta=factor(theta),
         cost_conditional=factor(cost_conditional),
         cost_other=factor(cost_likely)) %>%
  select(-cost_likely, -cost_neg, -cost_and)

params

p <- params %>% 
  ggplot(aes(x=alpha, y=MSE)) + geom_point(aes(color=cost_other, shape=cost_conditional)) +
  facet_grid(theta~cost_conditional, scales = "free")
p
ggsave(paste(target_dir, "MSE.png", sep=.Platform$file.sep), p,width=7,height=6)

p <- params %>% 
  ggplot(aes(x=alpha, y=MSE)) +
  geom_bar(aes(fill=theta), stat="identity", position="dodge") +
  facet_wrap(~cost_conditional, scales = "free") +
  theme(legend.position="bottom")
p

ggsave(paste(target_dir, "MSE-bars.png", sep=.Platform$file.sep), p, width=10, height=6)

# BEST n
n <-10
params %>% slice(1:n) %>% 
  # filter(MSE < 0.0025 - 0.0025/2) %>% 
  ggplot(aes(x=alpha, y=MSE)) +
  geom_point(aes(color=theta, shape=cost_conditional), size=2.5, alpha = 0.75) +
  ggtitle(paste("Best", n))
# WORST 10
params %>% slice(tail(row_number(), n)) %>% 
  # filter(MSE < 0.0025 - 0.0025/2) %>% 
  ggplot(aes(x=alpha, y=MSE)) + geom_point(aes(color=theta, shape=cost_conditional), size=2.5) +
  ggtitle(paste("Worst", n))

```

# optim

```{r, message=FALSE, warning=FALSE}
  # params <- read_csv(paste("../results/model/", now, "/parameters.csv", sep="")) %>%
  #   arrange(MSE)
params <- tibble(alpha=3, theta=0.7, cost_conditional=0.1, cost_other=10)
now <- str_replace_all(Sys.time(), "[\\s+:-]", "_")
target_dir <- paste(TARGET_DIR, "/optim/", now, "/", sep="")
dir.create(target_dir, showWarnings = FALSE, recursive=TRUE)
fit = optim(
    # initial parameter values
    par = c(params$alpha, params$theta, params$cost_conditional),
    # function to optimize
    fn = function(par) {
      dat.to_wppl$alpha <- par[1]
      dat.to_wppl$theta <- par[2]
      dat.to_wppl$cost_conditional <- par[3]
      dat.to_wppl$cost_and <- params$cost_other
      dat.to_wppl$cost_likely <- params$cost_other
      dat.to_wppl$cost_neg <- params$cost_other
      model_predictions <- run_model(dat.to_wppl)
      model <- model_predictions$avg
      # return(mse_model_human(dat.experiment2.means, model))
      return(-cor_model_human(dat.experiment2.means, model))
    },
    method="L-BFGS-B",
    lower=c(1, 0, 0, 0),
    upper=c(Inf, 0.95, Inf, Inf)
  )
fn <- paste(target_dir, "optim.csv", sep=.Platform$file.sep)
write.table(tibble(alpha=params$alpha, theta=params$theta,
                   cost_conditional=params$cost_conditional,
                   cost_other=params$cost_other,
                   optim_val=fit$value,
                   optim_alpha=fit$par[1], optim_theta=fit$par[2],
                   optim_cost_if=fit$par[3],
                   optim_cost_other=fit$par[4]),
            file = fn, row.names=FALSE, sep = ",", append = TRUE,
            col.names=!file.exists(fn))
write_rds(fit, paste(target_dir, "optim.rds", sep=.Platform$file.sep))

```

Collect params from all optim-runs

```{r, include=FALSE}
folders <- c("2020_02_15_09_06_47", "2020_02_15_08_05_01", "2020_02_15_00_15_21")

params <- tibble()
for(fn in folders){
  p <- read_csv(paste(TARGET_DIR, "optim", fn, "optim.csv", sep=.Platform$file.sep))
  params <- bind_rows(params, p)
}

```



